Argennon: A Scalable Cloud Based Smart Contract Platform Using Argument of Knowledge Systems

### 2022-09-17 tex/sections/consensus.tex
Update the normal consensus
We removed status blocks and an analysis is added for the effect of online/offline status in the protocol.

Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

The credibility of a block of the Argennon blockchain is determined by the certificates it has received
from different sets of users, known as committees. There are two primary type of certificate committees in
Argennon: the committee of \emph{delegates} and the assembly of \emph{validators}. Argennon has \emph{one} committee
of delegates and $m$ assemblies of validators.

The committee of delegates issues a certificate for every block of the Argennon blockchain, and each
assembly of validators issues a certificate every $m$ blocks. A validators assembly will
certify a block only if it has already been certified by the committee of delegates. Every assembly of validators has
an index between $0$ and $m - 1$, and it issues a certificate for block number $n$, if $n$ modulo $m$ equals
the assembly index.

@@

In addition to primary committees, Argennon has several community driven committees. Certificates of these
committees are not required for block finality, but they could be used by members of the
validators' committee to better decide about the validity of a block.

@@

The committee of delegates is a small committee of trusted delegates, elected by Argennon users through the
Argennon Decentralized Autonomous Governance system (ADAGs\footnote{pronounced \textipa{/eI-dagz/}.}).
At the start of the Argennon mainnet, this committee will have
five members, and later its size could be changed by the ADAGs in a procedure described
in Section~\ref{sec:adags}.

@@

Besides the main delegates' committee, a reserve committee of delegates consisting of three members is elected by users
either through the ADAGs or by \emph{emergency agreement} during the recovery protocol. In case the main committee
fails to generate new blocks or behaves maliciously, the task of
block generation will be assigned to the reserve committee until a new main delegates' committee is elected
through the ADAGs.

Usually, the delegates are large organizations, and they have enough computational resources to generate blocks
very fast. However, a block is not completely final if it does not have the certificate of its validators.
A certified block by the delegates will not be accepted by the network, if the last block certified by
the validators is behind it more than a certain number of blocks.

The committee of delegates may use any type of agreement protocol to reach consensus on the
next block. Usually a very simple and fast protocol can do the job: one of the members
is randomly chosen as the proposer, and other members vote "yes" or "no" on the proposed block.

If one of the delegates lose its network connectivity, no new blocks can be generated. For this reason,
the delegates should invest on different types of communication infrastructure, to make sure they never lose
connectivity to each other and to the Argennon network.

\subsection{Validators}\label{subsec:validators-committee}

The Argennon protocol calculates a stake value for every account, which is an estimate of a user's stake in the
system, and is measured in ARGs. Any account whose stake value is higher than
\texttt{minValidatorsStake} threshold is considered a \emph{validator}.
The \texttt{minValidatorsStake}
threshold is determined by the ADAGs, but it can never be higher than $1000$ ARGs.

Every \texttt{AssemblyLifeTime} number of blocks, randomly $m$ assemblies are selected from
validators, in a way that the total stakes of assemblies are almost equal, and every
account is a member of \textbf{at least} one assembly.

Every validator has a status which can be either \texttt{online} or \texttt{offline}.
This status is stored in the ARG smart contract and is a part of the staking database. A validator can change
his status through a method invocation
from the ARG smart contract. When an account sets its status to \texttt{offline}, it receives a small reward, and
it can not change it back to \texttt{online} for \texttt{statusCoolDown} number of blocks.

\note{When a validator changes his status, the change has no effect until the block containing the status
change transaction gets certified by his assembly.}

A block certificate issued by some members of a validators assembly is considered valid, if according to
the staking database of the previous block \textbf{certified by the same assembly}, we have:\footnote{If we calculate
the stake values based on the previous block a malicious assembly can select the validators of the next block.}
\begin{itemize}
    \item The total stake of \texttt{online} members of the assembly is higher than \texttt{minOnlineStake} fraction
    of the total stake of the assembly. This threshold can be changed by the ADAGs, but it can never be lower
    than $2/3$.
    \item All signers of the certificate have \texttt{online} status.
    \item The sum of stake values of the certificate signers is higher than $3/4$ of the total stake
    of the assembly members that have \texttt{online} status.
\end{itemize}

The delegates can generate blocks very fast. Therefore, the Argennon blockchain always has an
unvalidated part, which contains blocks that have a certificate from the assembly of delegates but have not
yet received a certificate from the validators.

As we mentioned before, the block with height $n$ needs a certificate from the assembly of
validators with index $n$ modulo $m$. To decide about signing the certificate of a block which already has
a certificate from the delegates, a validator checks the conditional
validity\footnote{See Section~\ref{subsec:block-validation}} of the block, and
if the block is valid he issues
an "accept" signature. If the block is invalid, he initiates the recovery protocol. The validator will broadcast the
certificate \textbf{only after} he sees the certificate of the validators of the previous block.
Some validators may also require seeing a certificate from
some community driven committees. An honest validator never signs a certificate for two different blocks with the
same height.

So in Argennon, the block validation by assemblies is performed in parallel, and validators
do not wait for seeing the certificate of the previous block assembly to start block validation. On the
other hand, the block certificates are published and broadcast sequentially. A validator does not publish its
certificate if the certificate of the previous block validator assembly has not been published yet. This ensures that
an invalid fork made by malicious delegates will not receive any certificates from any validator assemblies.

The value of $m$ is determined by the ADAGs. but it can never be higher than $25$. This way, it is guaranteed
that on average, any block of the Argennon blockchain is validated by at least $0.02$ of the total ARG supply.

block certificates issued by validators assemblies are included in the blocks of the Argennon blockchain.
A block can contain multiple certificates, provided that
those certificates belong to consecutive blocks.

\subsection{Status Blocks}\label{subsec:status-blocks}

If according to the staking database of block $n$, the total online stake of the committee with index $n$ modulo $m$ is
lower than \texttt{minOnlineStake} threshold, the block $n + m$ can never be certified by validators.

To prevent the blockchain from halting in such situations, the protocol performs a predefined partial
reshuffling of committee members.
In this reshuffling which is based on the block random seed, some online members from other committees will
be moved to the committee without enough online stake to make it active again.

If the reshuffling can not solve the problem due to low total online stake, the protocol requires the next block
of the blockchain to be a special \emph{status block}. A status block is a special block which can only contain
status change transactions. The status block needs to be certified by the delegates and by 2/3 of the total stake of
validators. The \texttt{online}/\texttt{offline} status of validators will not be considered in the validity of
the status block certificate.\footnote{Theoretically at the status block, the total online stake of the system could
be very low. Therefore, the status block should not be certified only by online stake.} After applying
the transactions of the status block, the total online stake of all committees of validators
must go higher than \texttt{minOnlineStake} threshold.

\subsection{Signature Aggregation}\label{subsec:sig-agg}

In Argennon, signature aggregation is mostly performed by PVC servers. To distribute the aggregation workload
between different servers, every validators assembly is divided into pre-determined groups, and each PVC
server is responsible for signature aggregation of one group. To make sure that there is enough redundancy, the
total number of groups should be less than the number of PVC servers and each group should be assigned to
multiple PVC servers.

Any member of a group knows all the servers that are responsible for signature aggregation of his group. When a member
signs a block certificate, he sends his signature to all of the severs that aggregate the signatures of his group.
These servers aggregate the signatures they receive and then send the aggregated signature to the delegates.
Furthermore, the delegates aggregate these signatures to produce the final block certificate
and then include it in the next block.

The role of the delegates in the signature aggregation is limited. The important part of the work is done by PVC
servers. As long as there are enough honest PVC servers, the network will be able to perform signature aggregation
even if the delegates are malicious.

>>>>

The credibility of a block of the Argennon blockchain is determined by the certificates it receives
from different sets of users, known as committees. There are two primary types of certificate committee in
Argennon: the committee of \emph{delegates} and the assembly of \emph{validators}. Argennon has \emph{one} committee
of delegates and $m$ assemblies of validators.

The committee of delegates issues a certificate for every block of the Argennon blockchain, and each
assembly of validators issues a certificate every $m$ blocks. A validator assembly will
certify a block only if it has already been certified by the committee of delegates. Every assembly of validators has
an index between $0$ and $m - 1$, and it issues a certificate for block with height $h$, if $h$ modulo $m$ equals
the assembly index.

@@

In addition to primary committees, Argennon has several community driven committees. Certificates of these
committees are not required for block finality, but they could be used by members of validator assemblies to better
decide about the validity of a block.

@@

The committee of delegates is a small committee of trusted delegates, elected by Argennon users through the
Argennon Decentralized Autonomous Governance system (ADAGs\footnote{pronounced \textipa{/eI-dagz/}.}).
At the start of the Argennon main-net, this committee will be elected for one-year terms and will have five members.
Later, this could be changed by the ADAGs in a procedure described in Section~\ref{sec:adags}.

@@

Besides the main committee, a reserve committee of delegates consisting of three members is elected by users
either through the ADAGs or by \emph{emergency agreement} during the recovery protocol. In case the main committee
fails to generate new blocks or behaves maliciously, the task of
block generation will be assigned to the reserve committee until the main committee comes back online or a new
committee is elected through the ADAGs.

A block certified only by the delegates committee is credible, but is not considered completely final until it
receives the certificate of its validator assembly. Since a block at height $h$ contains the validators certificate
of the block at height $h-k$, the unvalidated part of the blockchain can not be longer than $k$ blocks.

The committee of delegates may use any type of agreement protocol to reach consensus on the
next block. Usually, the delegates are large organizations, and they can communicate with each
other efficiently using their reliable networking infrastructure. This mostly eliminates the complications of their
consensus protocol and any protocol will have a good performance in practice. Usually a very simple and fast protocol
can do the job: one of the members is randomly chosen as the proposer, and other members vote "yes" or "no" on the
proposed block. For better performance, they should run their agreement protocol for reaching consensus about small
batches of transactions in their mem-pools, instead of the whole next block.

If one of the delegates loses its network connectivity, no new blocks can be generated. For this reason,
the delegates should invest on different types of communication infrastructure, to make sure they never lose
connectivity to each other and to the Argennon network.

\subsection{The Assemblies of Validators}\label{subsec:validators-committee}

The Argennon protocol calculates a stake value for every account, which is an estimate of a user's stake in the
system, and is measured in ARGs. Any account whose stake value is higher than
\texttt{minValidatorStake} threshold is considered a \emph{validator}.
The \texttt{minValidatorStake}
threshold is determined by the ADAGs, but it can never be higher than $1000$ ARGs.

Every \texttt{AssemblyLifeTime} number of blocks, randomly $m$ assemblies are selected from
validators, in a way that the total stakes of assemblies are almost equal, and every
account is a member of \textbf{at least} one assembly.\footnote{An account can be the member of multiple assemblies.}

The value of $m$ is determined by the ADAGs, but it can never be higher than $32$. This way, it is guaranteed
that on average, any block of the Argennon blockchain is validated by at least $3\%$ of the total ARG supply.

\subsubsection{Activity Status}

Every validator has a status which can be either \texttt{online} or \texttt{offline}.
This status is stored in the ARG smart contract and is part of the staking database. A validator can change
his status to \texttt{offline} through an external request (transaction) to the ARG application. In this request he
exactly specifies for how long he wants to be offline and after this period his status will automatically be considered
\texttt{online} again. When a validator sets his status to \texttt{offline} for some period of time, he
will receive a small portion of the maximum possible reward that a validator can receive in that period of time by
actively participating in the consensus protocol. This ensures that a validator has an incentive for changing his
status to offline rather than simply becoming inactive.

The staking database of a validator assembly can only be updated by the assembly itself. That means, an external
request which changes the status of a validator can only be included in a block that is being validated by the
assembly of that validator.

There is no transaction type for changing the status of a validator to \texttt{online}. A malicious delegates committee
would be able to censor this type of transactions and prevent honest validators from coming back online. For this reason
the status of a validator is considered online automatically when the specified period of time for being offline ends.

A block certificate issued by some members of a validators assembly is considered valid, if according to
the staking database of the previous block \textbf{certified by the same assembly}, we have:\footnote{If we calculate
the stake values based on the previous block, a malicious assembly can select the validators of the next block.}
\begin{itemize}
    \item The total stake of \texttt{online} members of the assembly is higher than \texttt{minOnlineStake} fraction
    of the total stake of the assembly. This threshold can be changed by the ADAGs, but it can never be lower
    than $2/3$.
    \item All signers of the certificate have \texttt{online} status.
    \item The sum of stake values of the certificate signers is higher than $3/4$ of the total stake
    of the assembly members that have \texttt{online} status.
\end{itemize}

If according to the staking database of block $h$, the total online stake of the assembly with index $h$ modulo $m$ is
lower than \texttt{minOnlineStake} threshold, the block $h + m$ can never be certified by validators. To prevent the
blockchain from halting in such situations, the validator assembly with index $h$ modulo $m$ will get merged into the
assembly that has the most online stake at block $h$. This will decrease the number of assemblies to $m-1$, and the
indices of assemblies will be updated appropriately.

\subsubsection{Certificate Signing}

The delegates can generate blocks very fast. Therefore, the Argennon blockchain always has an
unvalidated part which contains the blocks that have a certificate from the committee of delegates but have not
yet received a certificate from the validators.

As we mentioned before, the block with height $h$ needs a certificate from the assembly of
validators with index $h$ modulo $m$. To decide about signing the certificate of a block which already has
a certificate from the delegates, a validator checks the conditional
validity of the block (See Section~\ref{subsec:block-validation}), and if the block is valid, he issues
an "accept" signature. If the block is invalid, he initiates the recovery protocol. The validator will broadcast the
certificate \textbf{only after} he sees the certificate of the validator assembly of the previous block.
Some validators may also require seeing a certificate from
some community based committees. An honest validator never signs a certificate for two different blocks with the
same height.

Consequently, in Argennon the block validation by assemblies is performed in parallel, and validators
do not wait for seeing the validators certificate of the previous block to start block validation. On the
other hand, the block certificates are published and broadcast sequentially. A validator does not publish his
signature, if the certificate of the previous block validator assembly has not been published yet. This ensures that
an invalid fork made by malicious delegates will not receive any certificates from any validator assemblies.

\subsubsection{Analysis}

We analyze the minimum amount of stake that is required for conducting different types of attacks against the Argennon
blockchain. In these attack scenarios, we assume that a single validator assembly is corrupted, all the delegates
are malicious and the adversary is able to partition the network arbitrarily.

We denote the total stake of the corrupted validator assembly with $s$ and the total stake of malicious users of the
assembly with $m$. We use $d$ to denote the stake of the assembly's users with \texttt{offline} status and $h$ to denote
the total stake of the assembly's users with \texttt{online} status that do not participate in the
protocol.\footnote{$d$ stands for \emph{deactivated} and $h$ stands for \emph{hidden}.} We assume
that a certificate is accepted if it is signed by more than $r$ fraction of the total online stake of the assembly.
We obtain the minimum required malicious stake for three types of attacks:
\begin{itemize}
    \item Confirming an invalid block:
    \[ m > r(s-d) \]
    \item Forking the blockchain by double voting and network partitioning:
    \[m > (2r-1)(s-d)+h \]
    \item Halting the blockchain by refusing to vote:
    \[m > (1-r)(s-d)-h \]
\end{itemize}

In Argennon we have $r=\frac{3}{4}$ and $d < \frac{1}{3}s$.
Consequently, in Argennon to confirm an invalid block, the adversary needs at least $\frac{1}{2}$ of the total stake
of the assembly. For forking the blockchain, interestingly $m$ is
minimized when $h=0$, and $\frac{1}{3}s$ is the minimum required stake. For halting the blockchain,
an adversary requires a stake higher than $\frac{1}{6}s-h$.

In particular we are interested in comparing the Argennon protocol with a simple protocol that accepts a certificate
if it is signed by more than $\frac{2}{3}$ of the total stake of the assembly and there is no \texttt{online/offline}
status for users.

To compare the minimum stake required for halting the blockchain in two protocols, we can solve the following
inequality:
\[
    \frac{1}{4}(s-d)-h > \frac{1}{3}s-h\ \cdot
\]
We observe that as long as $d>\frac{1}{9}s$, the minimum required stake for halting the blockchain is higher in the
Argennon protocol.

\subsubsection{Signature Aggregation}\label{subsubsec:sig-agg}

The validators certificate of a block is an aggregate signature of members of the corresponding assembly of
validators. Validator assemblies could include millions of users and calculating their aggregate
signature requires an efficient distributed algorithm.

In Argennon, signature aggregation is mostly performed by PVC servers. To distribute the aggregation workload
between different servers, every validator assembly is divided into pre-determined groups, and each PVC
server is responsible for signature aggregation of one group. To make sure that there is enough redundancy, the
total number of groups should be less than the number of PVC servers and each group should be assigned to
multiple PVC servers.

Any member of a group knows all the servers that are responsible for signature aggregation of his group. When a member
signs a block certificate, he sends his signature to all the severs that aggregate the signatures of his group.
These PVC servers aggregate the signatures they receive and then send the aggregated signature to the delegates.
Furthermore, the delegates aggregate these signatures to produce the final block certificate
and then broadcast it to the PVC servers network.

The role of the delegates in the signature aggregation algorithm is limited. The important part of the work is done by
PVC servers and slightly modified versions of this algorithm can perform signature aggregation
even if all the delegates are malicious, as long as there are enough honest PVC servers.
